---
title: "Testosterone, diversity, and group project performance project"
author: "Cathy Su"
date: "9/7/2019"
output: pdf_document
---


## Assignment description

See: 
http://rosmarus.refsmmat.com/datasets/datasets/hormone-diversity/ 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1234)
# libraries
library(tidyverse)
library(lattice)
library(dagitty)
library(reshape2)
library(grid)
library(gridExtra)
# data
ind_dat <- read.csv("../data/Individual_Faultlines.csv",
                     sep = ",", 
                     header = T)
team_dat <- read.csv("../data/Team_Faultlines.csv",
                     sep = ",", 
                     header = T)
# make it a factor
ind_dat$team.id <- as.factor(ind_dat$team.id)
team_dat$team.id <- as.factor(team_dat$team.id)

theme_hw <- theme(plot.title = element_text(hjust = 0.5),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
       # panel.grid.major =   element_line(colour = "gray",size=0.5)
  )
```

## Introduction and data summary

```{r intro}
summary(ind_dat)
summary(team_dat)
# seems like interim.* columns contain a lot of missing data.

```

## Questions

### Q1

For each group, calculate the number of unique gender-ethnicity-country combinations (such as female-white-Russia or male-Indian-USA) among the group members, and store this with the other group information such as team size and performance. Also calculate the average testosterone level for each group.

```{r score, fig.width=6, fig.height=3}
get_score <-function(group){
  score <- length(unique(ind_dat$combo[ind_dat$team.id == group,]))
  return(score)
}
#  calculate the number of unique gender-ethnicity-country combinations
ind_dat$combo <-paste(ind_dat$Gender, ind_dat$Ethnicity, ind_dat$Country)
team_dat$score<- unlist(lapply(team_dat$team.id, 
               function(x){length(unique(ind_dat$combo[ind_dat$team.id == x]))}))

#  calculate the average testosterone level for each group.
team_dat$avg.log.testosterone<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.testosterone[ind_dat$team.id == x])}))

#  calculate the average cortisol level for each group.
team_dat$avg.log.cortisol<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.cortisol[ind_dat$team.id == x])}))

```

### Q2

Do exploratory data analysis to explore the composition of groups, the typical amount of diversity, and the typical amounts of testosterone. Note particularly that the data includes the logs of the cortisol and testosterone levels as well as the raw levels; does your EDA suggest you should use the logs or the raw values?

#### Composition of groups 

```{r groups, fig.width=5, fig.height=2}

# visualise the distribution of diversity
ggplot(team_dat, aes(x = score)) +
  geom_histogram(bins = 4)+ labs(title="Gender-ethnicity-country score")

# visualise the distribution of testosterone
ggplot(team_dat, aes(x = avg.log.testosterone)) +
  geom_histogram(bins = 4)+ labs(title="Avg log testosterone per team")

p1 <-ggplot(ind_dat, aes(x= team.id, y=Testosterone))+
  geom_boxplot()+theme_hw

p2 <-ggplot(ind_dat, aes(x= team.id, y=log.testosterone))+
  geom_boxplot()+theme_hw

grid.arrange(p1, p2, ncol = 2)

# visualise the distribution of cortisol
ggplot(team_dat, aes(x = avg.log.cortisol)) +
  geom_histogram(bins = 4)+ labs(title="Avg log cortisol per team")

p1 <-ggplot(ind_dat, aes(x= team.id, y=Cortisol))+
  geom_boxplot()+theme_hw

p2 <-ggplot(ind_dat, aes(x= team.id, y=log.cortisol))+
  geom_boxplot()+theme_hw

grid.arrange(p1, p2, ncol = 2)
```

### Sketch out causal diagrams

The findings suggest that diversity is beneficial for performance, but only if group-level testosterone is low; diversity has a negative effect on performance if group-level testosterone is high.

```{r causal, fig.width=3, fig.height=2}
library(dagitty)

g <- dagitty('dag {
    testosterone  [pos="0,0"]
    diversity [pos="0,1"]
    performance [pos="1,0.5"]
    cortisol [pos="2,0"]
    
    testosterone -> performance <- diversity
    cortisol-> performance 
}')
plot(g)

impliedConditionalIndependencies( g )
```


### Q3

### Q4

### Q5

