---
title: "Testosterone, diversity, and group project performance"
author: "Cathy Su"
date: "27/10/2019"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
bibliography: hw3.bib
---
```{r global_options, include=FALSE}
# knitr::knit_hooks$set(plot = function(x, options)  {
#   hook_plot_tex(x, options)
# })
library(knitr)
opts_knit$set(eval.after = 'fig.cap')
```

```{r setup, include=FALSE}
#http://rosmarus.refsmmat.com/datasets/datasets/hormone-diversity/ 
  
knitr::opts_chunk$set(echo = FALSE,  warning=FALSE,width.cutoff=60)
# , fig.pos = 'H'
set.seed(1234)

mod_stargazer <- function(...){
  output <- capture.output(stargazer(...))
  # The first three lines are the ones we want to remove...
  output <- output[4:length(output)]
  # cat out the results - this is essentially just what stargazer does too
  cat(paste(output, collapse = "\n"), "\n")
}
##### example EDA
# https://rstudio-pubs-static.s3.amazonaws.com/298962_d4d5811f31394f04b707ce56b43b74c5.html
# 
# https://rdrr.io/cran/dlookr/f/vignettes/EDA.Rmd
# 
# https://github.com/jrnold/r4ds-exercise-solutions/blob/master/EDA.Rmd
# 
# https://github.com/rdpeng/artofdatascience/blob/master/manuscript/EDA.Rmd

# https://github.com/Xiaodan/Coursera-Regression-Models/blob/master/motor_trend_project/report.Rmd

# http://jgscott.github.io/teaching/writeups/files/example_writeup1.pdf

# libraries
library(stargazer)
library(tidyverse)
library(lattice)
library(dagitty)
library(reshape2)
library(grid)
library(gridExtra)
library("ggpubr")
library(knitr)
library("sjPlot")
# data
ind_dat <- read.csv("../data/Individual_Faultlines.csv",
                     sep = ",", 
                     header = T)
team_dat <- read.csv("../data/Team_Faultlines.csv",
                     sep = ",", 
                     header = T)
# make it a factor
ind_dat$team.id <- as.factor(ind_dat$team.id)
team_dat$team.id <- as.factor(team_dat$team.id)

theme_hw <- theme(plot.title = element_text(hjust = 0.5),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
       # panel.grid.major =   element_line(colour = "gray",size=0.5)
  )

#  calculate the number of unique gender-ethnicity-country combinations
ind_dat$combo <-paste(ind_dat$Gender, ind_dat$Ethnicity, ind_dat$Country)
team_dat$diversity.score<- unlist(lapply(team_dat$team.id, 
               function(x){length(unique(ind_dat$combo[ind_dat$team.id == x]))}))
#team_dat$diversity.score <- team_dat$score/team_dat$team.size #suggested by Q4
team_dat$proportion.females <- team_dat$females/team_dat$team.size
#  calculate the average testosterone level for each group.
# some have missing testosterone data, this means we need to average. 
team_dat$avg.log.testosterone<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.testosterone[(ind_dat$team.id == x)], na.rm = TRUE)}))

#  calculate the average cortisol level for each group.
team_dat$avg.log.cortisol<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.cortisol[ind_dat$team.id == x], na.rm = TRUE)}))

#  calculate the average age for each group.
team_dat$avg.age<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$Age[ind_dat$team.id == x], na.rm = TRUE)}))

team_dat$age.variance<- unlist(lapply(team_dat$team.id, 
               function(x){var(ind_dat$Age[ind_dat$team.id == x], na.rm = TRUE)}))
```

## Executive Summary

In this report, we analyze a demographic data set collected by *[@Akinola2018]* and explore the relationship between a set of variables that contribute to group performance on a competetive task. The data comprises individual level and group level statistics collected from groups of MBA students completing a 7-day group project. We use  exploratory data analysis and regression models to mainly explore how **diversity, testosterone and cortisol** levels affect **final.performance**. We fit several linear regression models and found that when not accounting for cortisol, indeed diversity has a positive effect on performance, but only if group-level testosterone is low. If we account for cortisol and not testosterone, diversity has a positive effect on performance, but only if group-level cortisol is low. Additionally both hormones alone have a positive effect on group performance suggesting both stress and testosterone contribute to performance. To resolve the results of these two different models, we then used cross validation to select the model with the best performance by best subset regression, which contains both hormones as well as their interaction effects with diversity. 

## Introduction

Diversity and conflict are considered important factors which influence how well we work in groups [@vanK]. As the working world becomes more connected across the globe and thus the diversity of organizational groups increases, it is important to characterize the effect of diversity on group performance. Previous work by [@Akinola2018] suggests that both diversity and group hormone levels will influence how well groups perform on a competetive task. In their study, they considered levels of the two hormones testosterone and cortisol. Testosterone is involved in dominance and competition related behaviour in individuals and is produced at a higher level in males than females, while cortisol is a hormone released during physical and psychological stress [@MEHTA2015163]. For healthy males between 19 to 40 years, normal testosterone is known to fall within the 15.4 to 13 nmol/L range [@Kelsey2014].

In their work, [@Akinola2018] collected both demographic data and hormone measurements from 370 MBA students organized into 74 groups who partcipated in a competitive week long project where their goal was to outperform other groups. There were 370 individuals randomly organized into 74 groups.  Based on their demographic and hormone measurement data, the authors concluded that diversity is beneficial for performance, but only if group-level testosterone is low; and diversity has a negative effect on performance if group-level testosterone is high. However, the authors did not mention analyzing cortisol even though cortisol levels is suggested to have an effect testosterone's role in status-relevant behavior [@MEHTA2015163].

To validate the author's hypothesis and additionally examine the specific role of cortisol, we are using the [@Akinola2018] dataset which has been processed by [Nifty Datasets](http://rosmarus.refsmmat.com/datasets/datasets/hormone-diversity/) into separate individual level and group level datasets. Based on the preamble, we hypothesized that the effects of testosterone and diversity on performance are mediated by their opposite effects on 'cooperation' (not directly measured) in the group. Furthermore cortisol levels largely unevaluated by the study may influence performance through affecting group 'stress' (not directly measured). Putting this together with the measured variables, we illustrate our hypotheses about the data in Figure \ref{fig:cause}. Here 'interim.other' describes other interim measurements of group performance which were in the dataset and 'final.other' describes the measurements of group performance at the conclusion of the task which contribute to the **final.performance** score. This diagram helps set the context for reasoning about which regression models we should try.

```{r causal, fig.width=9, fig.height=2.5, echo=FALSE, fig.cap="\\label{fig:cause}Causal diagram illustrates hypothesized relationships of experimental variables involved in relationship between testosterone and final group performance."}

g <- dagitty('dag {
    testosterone [pos="0,0"]
    diversity [pos="0,1"]
    final.performance [pos="4,0.5"]
    final.other [pos="4,1"]
    interim.other[pos="3,1"]
    cortisol [pos="1.2,0"]
    cooperation [pos="2,0.5"]
    gender.ethnicity.country [pos="0,2"]
    team.females [pos="-1,1"]
    age [pos="-1,1.5"]
    team.size  [pos="-1,2.0"]
    stress [pos="2,0"]
    
    cortisol -> stress -> cooperation
    gender.ethnicity.country <- team.females ->testosterone
    team.females -> diversity <- team.size 
    diversity <- age ->testosterone
    cooperation -> final.performance <- final.other
    gender.ethnicity.country -> diversity -> cooperation 
    testosterone ->cooperation-> final.other
    cooperation-> interim.other-> final.other
}')

plot(g)
```

## Methods

### Calculation of group level variables and removal of outliers

We are interested in doing our analysis at the group level therefore we needed to aggregate the individual level data. We saw there were <10 individuals with partly missing data. Since we are trying to look at team level performance, we did not remove any individuals. For these individuals, not everything was missing so we calculated group average measurements, e.g. average hormone measurements, from other members.
Additionally, we have calculated group diversity score as the number of unique gender-ethnicity-country combinations present in the group. Lastly we calculate proportion of females in the group as the number of females divided by group size. 
We then examined whether there were any groups with outlier measurements for the key variables: diversity score, hormone measurements and team performance. 

## Exploratory Data Analysis & Data Summary

After aggregating data to the group level, we checked the distributions of all of the variables which is summarized below. Only measurements in the 'interim' variables are missing. Given that it's unclear how the multiple interim measurements may relate to the final score and they contain many missing values, we removed these variables. However, the limitation is that we may be very dependent upon a single measure of performance, which would be the score assigned to the variable "final.performance". 
```{r}
summary(team_dat)
```

### Distribution of hormone levels across individuals and groups

It was clear when for both hormone levels that the log transformed values were distributed with less skew across teams than the raw values and have fewer outlier values. This is preferable so we chose like the authors to use averaged log testosterone per group. Figure \ref{fig:test} shows the distribtuions for testosterone but for cortisol the difference is similar. 

```{r test, fig.width=8, fig.height=3, echo=FALSE, message=FALSE, warning= FALSE, fig.cap="\\label{fig:test}Distributions of testosterone and log testosterone levels in each team."}
p1 <-ggplot(ind_dat, aes(x= team.id, y=Testosterone))+
  geom_boxplot()+theme_hw + ylab("Testosterone in pg/mL")
p2 <-ggplot(ind_dat, aes(x= team.id, y=log.testosterone))+
  geom_boxplot()+theme_hw + ylab("Log(Testosterone in pg/mL)")
p3 <-ggplot(ind_dat, aes(x= team.id, y=Cortisol))+
  geom_boxplot()+theme_hw + ylab("Cortisol in nMol/L")
p4 <-ggplot(ind_dat, aes(x= team.id, y=log.cortisol))+
  geom_boxplot()+theme_hw + ylab("Log(cortisol in nMol/L)")
grid.arrange(p1, p2, 
             #p3, p4, 
             ncol = 2)
```

### Incorporation of age and gender 

Both age and gender were included in the results of the original manuscript. Based upon our causal graph, both can have an influence on final performance through influencing testosterone levels or through influencing diversity. We know we need to study their impact on teamwork, therefore it makes sense to look at the proportion of females in the group and the variance of age in the group which are measures of diversity. We also know that hormone levels depend upon age, therefore we also calculate average age in the group.  

### Univariate and pairwise distributions of group level variables

The univariate distributions of the group level variables is given across the diagonal in Figure \ref{fig:pairs}. We see that in particular, our diversity score appears bimodal. Although our score is calculated differently, [@Akinola2018] classified diversity score into two bins in their faultline analysis suggesting that our diversity score may behave similarly. 

In the same figure we have the pairwise comparisons of the important variables as well. In the upper diagonal, the Pearson correlation coefficients (upper right half) between important variables are described with their significance. 

Based upon the summary statistics and the distributions seen in Figure \ref{fig:pairs}, we noted the following outlier teams:

* we likely do not need to discard variables based on collinearity.
* teams 65, 92 and 101 had very low final performance (below a score of -2).
* team 64 had very low log testosterone.
* team 39 was the only one with no female members.

However, we did not remove these outliers at the outset because we actually don't have many teams to fit the model on and the outliers lay within wo standard deviations of the mean of each variable in question.

```{r dists, fig.width=10, fig.height=10, echo=FALSE, message=FALSE, warning= FALSE,fig.cap="\\label{fig:pairs}Pairwise correlations of important variables including their Pearson correlation coefficient. Significant correlations are marked by the corresponding number of astericks."}
# check relationships of all variables of interest
allvars <- colnames(team_dat)[c(2:4, 15:20)]
vars <- colnames(team_dat)[c(2:4, 15:18)]
#cor(team_dat)
#pairs(team_dat[vars], pch = 19,  lower.panel=NULL)
library("PerformanceAnalytics")
chart.Correlation(team_dat[allvars], histogram=TRUE, pch=19)

# removal of variables
#probs <- c(64, 65, 92, 101, 39)
#team_dat <- team_dat[!(team_dat$team.id %in% probs),]
```

## Results

### Selecting the variables with non-negligible effect

Out of the variables that we had considered in Figure \ref{fig:pairs}, to build models we first performed variable selection to exclude those without large effects. We included the interaction terms between diversity score and each of the hormones, which are interpretable interaction terms that we need to consider to answer the substantive questions. 
First we picked how many terms we should have in the best predictive linear model by using 10-fold cross validation and plotted the mean squared error in Figure \ref{fig:cv}. This analysis shows validation error is lowest around 8 terms.

```{r cv, echo=FALSE, message=FALSE, warning= FALSE, fig.cap="\\label{fig:cv} Cross validation error for each number of predictors"}

library(leaps)
set.seed(12)
folds=sample(rep(1:10,length=nrow(team_dat)))

predict.regsubsets=function(object,newdata,id,...){
  form=as.formula(object$call[[2]])
  mat=model.matrix(form,newdata)
  coefi=coef(object,id=id)
  mat[,names(coefi)]%*%coefi
}

vars <-c(allvars)
cv.errors=matrix(NA,10,11)
for(k in 1:10){
  # best fit on the train data
  best.fit=regsubsets(final.performance~. + avg.log.testosterone:diversity.score + diversity.score:avg.log.cortisol, data=team_dat[folds!=k, vars],nvmax=40,method="exhaustive")
  # performance on the test data
  for(i in 1:10){
    pred=predict(best.fit,team_dat[folds==k,vars],id=i)
    cv.errors[k,i]=mean((team_dat$final.performance[folds==k]-pred)^2)
  }
}

# Take the mean of over all folds for each model size
mean_cv_errors = apply(cv.errors, 2, mean)

# Find the model size with the smallest cross-validation error
min = which.min(mean_cv_errors)

# Plot the cross-validation error for each model size, highlight the min
plot(mean_cv_errors, type='b', xlab = "Number of predictors",
     ylab = "Mean cross validation error (percent)")
points(min, mean_cv_errors[min][1], col = "red", cex = 2, pch = 20)

```

Based on this we used best subset selection on the full data set in order to obtain the best  8-predictor model. The variables and coefficients selected by the model are given in Table 5.  While the coefficients of the predictive model are not necessarily useful for explaining the final group performance, the predictive model includes only variables with non-negligible effect. This suggests to us that age variance for example is not a useful explanatory variable, and we discarded these extra variables.

```{r, echo=FALSE}
regfit_best = regsubsets(final.performance~. + avg.log.testosterone:diversity.score +diversity.score:avg.log.cortisol, data = team_dat[,vars], nvmax = 8)
kable(coef(regfit_best, 8), format = "latex",caption = "Variables selected for 8-variable model found by best subsets regression")
```
### Effect of group diversity on relationship between testosterone and performance

The results discussed by the original study [@Akinola2018] suggest that when group diversity was low, group testosterone significantly positively predicted performance at p < .01  when group diversity was relatively high, group testosterone significantly negatively predicted performance p <0.01. However they were not considering cortisol at the time. Given this we fit a model of final performance using all of the variables chosen from Table 5 except those terms which included cortisol. The coefficients which we found are displayed in Table X. 

```{r test_model, fig.width=8, fig.height=8, warning=FALSE, results='asis',echo=FALSE, fig.cap="Model results with interaction between diversity and performance"}
# model with normalized score
#mod_base =lm(final.performance ~ avg.log.testosterone + diversity.score + team.size, data = team_dat)
#mod_int=lm(final.performance ~ avg.log.testosterone + diversity.score + team.size + avg.log.testosterone:diversity.score, data = team_dat)
#summary(mod3)
#kable(mod3$coef, digits = c(3,3,3,4), format = 'markdown')
#par(mfrow = c(2, 2)) 
#plot(mod, col=c("blue"))

#full_form1 <-as.formula(paste("final.performance ~ ", paste(vars, collapse= "+")))
form_sans_cortisol <- as.formula(paste("final.performance ~ ", paste(c(vars[1:6],"avg.log.testosterone:diversity.score"), collapse= "+")))
base <- as.formula(paste("final.performance ~ ", paste(c(vars[1:6], vars[8:9]),collapse= "+")))
full_mod1 <- lm(form_sans_cortisol, data = team_dat)
full_mod2<- lm( base, data = team_dat)
full_form2 <-as.formula(paste("final.performance ~ ", paste(vars, collapse= "+"), "+avg.log.testosterone:diversity.score",
                             "+avg.log.cortisol:diversity.score", sep = ""))
full_form3 <-as.formula(paste("final.performance ~ ", paste(vars, collapse= "+"), "+avg.log.testosterone:diversity.score", sep = ""))
full_form4 <-as.formula(paste("final.performance ~ ", paste(vars, collapse= "+"), "+avg.log.cortisol:diversity.score", sep = ""))

stargazer(full_mod1,
         # model,
        # object.names = TRUE,
         column.labels = c("model 1", "model 2"),
          model.names = FALSE,
        model.numbers = FALSE,
          flip = TRUE,
          header=FALSE, type = "latex",
          intercept.bottom = FALSE, out.header = FALSE,
          label = "tab:regression",
          font.size="small",
          column.sep.width = "1pt",
          single.row = TRUE,
          omit.table.layout = "n",
          #model.names = FALSE, # remove model names
          title = "Terms included in model of testosterone and performance")
```

We verified that this model could provide a good fit to the data by looking at the diagnostic plots. We found that there were two outliers with low residuals (marked 66 and 59) in the plot of residuals vs. fitted and the normal QQ-plot, but these were not highly influential points as measured by Cook's distance which is shown in the Residuals vs. Leverage plot.

```{r}
par(mfrow = c(2, 2)) 
plot(full_mod1)
```

In this model, the interaction term diversity.score:avg.log.testosterone is indeed significant (coefficient = -2.2574, p < 0.001; see Table 1). The negative sign of the coefficient implies that there is an opposite effect on performance from each of the two predictors diversity and testosterone. This is better illustrated in Figure \ref{fig:int} where we can see that when diversity is low at 3 units, group testosterone positively correlates with performance whereas when diversity is high at 6 units testosterone negatively correlates with performance. This suggests that we have verified the findings of the authors.

```{r, echo=FALSE,fig.cap="\\label{fig:int}Interaction between diversity and performance. The plot suggests that as the authors found, group diversity mediates teh effect of testosterone on performance."}

plot_model(full_mod1, type = "int") 
```

### Effect of cortisol on relationship between diversity and performance

Our EDA and variable selection had shown us that cortisol might be useful to include in our explanatory model of performance and so we additionally tested a model which includes cortisol levels and their interaction with diversity (Table X model X). However we found that in this model, neither cortisol nor its interaction with diversity are found to have a significant effect at p < 0.05. 

```{r, results='asis'}
form_sans_cortisol <- as.formula(paste("final.performance ~ ", paste(c(vars[1:6],"avg.log.testosterone:diversity.score"), collapse= "+")))

full_form4 <-as.formula(paste("final.performance ~ ", 
                              paste( c(vars, 
                                    "avg.log.cortisol:diversity.score",
                                    "avg.log.testosterone:diversity.score"),
                                    collapse= "+"), 
                              sep = ""))
model1 <- lm(full_form4, data = team_dat)
model2 <- lm(full_form2, data = team_dat)
stargazer(model1,
          model.names = FALSE,
           model.numbers = FALSE,
          #object.names = TRUE,
        column.labels = c("model 3", "model 4"),
          flip = TRUE,
          header=FALSE, type = "latex",
          intercept.bottom = FALSE, out.header = FALSE,
          label = "tab:regression",
          font.size="small",
          column.sep.width = "1pt",
          single.row = TRUE,
          omit.table.layout = "n",
          #model.names = FALSE, # remove model names
          title = "Terms included in model of cortisol, testosterone and performance")
```

Again we verified that this model could provide a good fit to the data by looking at the diagnostic plots. The same two outliers with low residuals found in the model of testosterone and performance were present here (marked 66 and 59) but these were these were not highly influential points as measured by Cook's distance which is shown in the Residuals vs. Leverage plot. We also note that the value of teh adjusted R squared in this model is significantly higher than in the model without cortisol terms.

```{r}
par(mfrow = c(2, 2)) 
plot(model1)
```

In this model, the average log cortisol variable has a positive coefficient suggesting that stressed groups have better performance. There is a negative coefficient on the diversity.score:avg.log.cortisol term suggesting that like with testosterone, stress changes the effect of diversity on performance and specifically a unit increase in average log cortisol has an antagonizing effect to a unit increase in diversity. However the terms are not found to be significant so we will need to do further study to determine whether cortisol has an effect on group performance or not.

### Conclusion

Here we have analyzed demographic data and hormone measurements from groups of MBA students performing a competetive project, previously published by [@Akinola2018]. We sought to investigate the authors' hypothesis that group diversity has a testosterone-dependent effect on group performance and also to check whether cortisol levels had an effect on this relationship.

By building linear models and comparing the nested models with an F-test, we have shown that when we do not account for cortisol the interaction between diversity and testosterone has a significant negative effect on performance (p < 0.01) implying that high diversity and high testosterone are antagonizing factors. This agrees with the authors' findings that diversity is beneficial for performance, but only if group-level testosterone is low.

Additionally, we found that when we incorporate terms for cortisol and its interaction with diversity without accounting for the interaction of testosterone with diversity, we build a linear model where stress has a positive effect on performance but stress and group diversity have antagonistic effects in interaction. Additionally both hormones and their interaction with diversity were found to be part of the best predictive model of performance by best subset regression. This analysis suggests that perhaps, stress has a role in group performance as well which merits further investigation.

Although we had some similar findings to the original study when examining diversity and testosterone, our results may not be directly comparable because of some differences in our methodology. Most prominently, [@Akinola2018] have used a faultline analysis to evaluate diversity whereas we have constructed a diversity score. As well, we have not included some of the variables that are present in the models which they tested e.g. proportion of females. We chose to discard these variables based upon our EDA and our reasoning about the relationship between variables collected in the study. Lastly we cannot compare our findings about cortisol because this was not discussed in depth in their original analysis. However our analysis suggests that although cortisol is useful to measure when looking at group performance, the measurements of cortisol and testosterone may be confounded.

# Bibliography
