---
title: "Testosterone, diversity, and group project performance"
author: "Cathy Su"
date: "9/10/2019"
output: 
  pdf_document:
    highlight: tango
    fig_caption: yes
bibliography: hw3.bib
---

```{r setup, include=FALSE}
#http://rosmarus.refsmmat.com/datasets/datasets/hormone-diversity/ 
  
knitr::opts_chunk$set(echo = FALSE, width.cutoff=60, fig.pos = 'H')
set.seed(1234)

##### example EDA
# https://rstudio-pubs-static.s3.amazonaws.com/298962_d4d5811f31394f04b707ce56b43b74c5.html
# 
# https://rdrr.io/cran/dlookr/f/vignettes/EDA.Rmd
# 
# https://github.com/jrnold/r4ds-exercise-solutions/blob/master/EDA.Rmd
# 
# https://github.com/rdpeng/artofdatascience/blob/master/manuscript/EDA.Rmd

# https://github.com/Xiaodan/Coursera-Regression-Models/blob/master/motor_trend_project/report.Rmd

# http://jgscott.github.io/teaching/writeups/files/example_writeup1.pdf

# libraries
library(tidyverse)
library(lattice)
library(dagitty)
library(reshape2)
library(grid)
library(gridExtra)
library("ggpubr")
# data
ind_dat <- read.csv("../data/Individual_Faultlines.csv",
                     sep = ",", 
                     header = T)
team_dat <- read.csv("../data/Team_Faultlines.csv",
                     sep = ",", 
                     header = T)
# make it a factor
ind_dat$team.id <- as.factor(ind_dat$team.id)
team_dat$team.id <- as.factor(team_dat$team.id)

theme_hw <- theme(plot.title = element_text(hjust = 0.5),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
       # panel.grid.major =   element_line(colour = "gray",size=0.5)
  )

```

## Executive Summary

In this report, we analyzed `mtcars` data set and explore the relationship between a set of variables and miles per gallon (MPG). The data was extracted from the paper *[@Akinola2018]* and comprises individual level and group level statistics collected from groups of MBA students who partcipated in a competitive 7-day group project. We use regression models and exploratory data analyses to mainly explore how **diversity, cortisol** and **testosterone** levels affect the final **team performance** score. The t-test shows that the performance difference between cars with automatic and manual transmission. And it is about 7 MPG more for cars with manual transmission than those with automatic transmission. 
Then, we fit several linear regression models and select the one with highest Adjusted R-squared value. So, given that weight and 1/4 mile time are held constant, manual transmitted cars are 14.079 + (-4.141)*weight more MPG (miles per gallon) on average better than automatic transmitted cars. Thus, cars that are lighter in weight with a manual transmission and cars that are heavier in weight with an automatic transmission will have higher MPG values.


## Introduction

Diversity and conflict are considered important factors which influence how well we work in groups [@vanK]. As the working world becomes more connected across the globe and thus the diversity of organizational groups increases, it is important to characterize the effect of diversity on group performance. Previous work by [@Akinola2018] suggests that both diversity and group hormone levels will influence how well groups perform on a competetive task. In their study, they considered levels of the two hormones testosterone and cortisol. Testosterone is involved in dominance and competition related behaviour in individuals and is produced at a higher level in males than females, while cortisol is a hormone released during physical and psychological stress [@MEHTA2015163].  

In their work, [@Akinola2018] collected both demographic data and hormone measurements from 370 MBA students organized into 74 groups who partcipated in a competitive week long project where their goal was to outperform other groups. There were 370 individuals randomly organized into 74 groups.  Based on their demographic and hormone measurement data, the authors concluded that diversity is beneficial for performance, but only if group-level testosterone is low; and diversity has a negative effect on performance if group-level testosterone is high. However, the authors did not mention analyzing cortisol even though cortisol levels is suggested to have an effect testosterone's role in status-relevant behavior [@MEHTA2015163].

To validate the author's hypothesis and examine the specific role of cortisol, we have obtained the [@Akinola2018] dataset which has been processed by [Nifty Datasets](http://rosmarus.refsmmat.com/datasets/datasets/hormone-diversity/) into separate individual level and group level datasets. Here we test the interactions between the hormone profiles of both cortisol and testosterone by modelling their effect on performance in the context of the demographic variables collected and the group diversity.

### Causal diagram 

Based on the preamble from [@Akinola2018] we may guess that the effects of testosterone and diversity on performance are mediated by their opposite effects on 'cooperation' (not directly measured) in the group. Furthermore cortisol levels largely unevaluated by the study may influence performance through affecting group 'stress' (not directly measured). Putting this together with the measured variables, our hypothesized causal diagram follows:

```{r causal, fig.width=8, fig.height=2.5, echo=FALSE, fig.cap="\\label{fig:figs}Causal diagram of experimental variables"}

g <- dagitty('dag {
    testosterone [pos="0,0"]
    diversity [pos="0,1"]
    final.performance [pos="4,0.5"]
    final.other [pos="4,1"]
    interim.other[pos="3,1"]
    cortisol [pos="1.2,0"]
    cooperation [pos="2,0.5"]
    gender.ethnicity.country [pos="0,2"]
    team.females [pos="-1,1"]
    age [pos="-1,1.5"]
    team.size  [pos="-1,2.0"]
    stress [pos="2,0"]
    
    cortisol -> stress -> cooperation
    gender.ethnicity.country <- team.females ->testosterone
    team.females -> diversity <- team.size 
    diversity <- age ->testosterone
    cooperation -> final.performance <- final.other
    gender.ethnicity.country -> diversity -> cooperation 
    testosterone ->cooperation-> final.other
    cooperation-> interim.other-> final.other
}')

plot(g)
```

## Methods

### Handling missing data 
Before calculating additional team level statistics, we saw there were <10 individuals with partly missing data (shown below). We preferred to not remove any individuals since we are trying to look at team level performance. For these individuals, not everything was missing so we calculated group average measurements, e.g. average hormone measurements, from other members. When we did this, there were no teams with missing data aside from in the 'interim' variables.

```{r missing, fig.width=8, fig.height=2, warning= FALSE}
head(ind_dat[!complete.cases(ind_dat),])
```

### Calculation of group level variables from individual level variables

We are interested in doing our analysis at the group level therefore we needed to calculate the group level values for the diversity score, average log testosterone level, average log cortisol level, average age and average age variance. Note that unlike in the original study, we have calculated group diversity score as the number of unique gender-ethnicity-country combinations (normalized by group size).
For the sake of making the units clear we show the code below. With this we can proceed to data exploration.

```{r groups, fig.width=8, fig.height=2, warning= FALSE, echo= TRUE, fig.cap="\\label{fig:figs}Causal Diagram"}
#  calculate the number of unique gender-ethnicity-country combinations
ind_dat$combo <-paste(ind_dat$Gender, ind_dat$Ethnicity, ind_dat$Country)
team_dat$score<- unlist(lapply(team_dat$team.id, 
               function(x){length(unique(ind_dat$combo[ind_dat$team.id == x]))}))
team_dat$diversity.score <- team_dat$score/team_dat$team.size #suggested by Q4
team_dat$proportion.females <- team_dat$females/team_dat$team.size
#  calculate the average testosterone level for each group.
# some have missing testosterone data, this means we need to average. 
team_dat$avg.log.testosterone<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.testosterone[(ind_dat$team.id == x)], na.rm = TRUE)}))

#  calculate the average cortisol level for each group.
team_dat$avg.log.cortisol<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$log.cortisol[ind_dat$team.id == x], na.rm = TRUE)}))

#  calculate the average age for each group.
team_dat$avg.age<- unlist(lapply(team_dat$team.id, 
               function(x){mean(ind_dat$Age[ind_dat$team.id == x], na.rm = TRUE)}))

team_dat$age.variance<- unlist(lapply(team_dat$team.id, 
               function(x){var(ind_dat$Age[ind_dat$team.id == x], na.rm = TRUE)}))

```

### Log vs raw values of testosterone and cortisol

Since we are interested in how the the levels of these two hormones may be related to the other variables, we first wanted to check whether we should use their log vs raw values in our EDA.

```{r test, fig.width=8, fig.height=3, echo=FALSE, warning= FALSE}
p1 <-ggplot(ind_dat, aes(x= team.id, y=Testosterone))+
  geom_boxplot()+theme_hw
p2 <-ggplot(ind_dat, aes(x= team.id, y=log.testosterone))+
  geom_boxplot()+theme_hw
grid.arrange(p1, p2, ncol = 2,
            top =textGrob("Testosterone vs. log testosterone levels in each team" ,
                          gp=gpar(fontsize=12,font=3)))# testosterone looks similar, except for scale.
```

(Cortisol looks similar to the above, except for scale.) It was clear when for both hormones that the log transformed values were distributed with less skew across teams which would be our preference. The authors also chose to use averaged log testosterone per group.

## Exploratory Data Analysis & Data Summary

### Distributions of variables

We checked the distribution of other key variables besides hormone level in our model, and saw that in particular, our diversity score and team size appear bimodal. Although our score is differently calculated, this agrees with [@Akinola2018] which classified diversity score into two bins.

```{r dists, fig.width=8, fig.height=2, echo=FALSE, warning= FALSE}
p1 <-ggplot(team_dat, aes(x=score))+
  geom_density()
p2 <-ggplot(team_dat, aes(x=proportion.females))+
  geom_density()
p3 <-ggplot(team_dat, aes(x=avg.age))+
  geom_density()
p4 <-ggplot(team_dat, aes(x=team.size))+
  geom_density()
p5 <-ggplot(team_dat, aes(x = team.size, y = score)) +
  geom_count() +theme(plot.title = element_text(hjust = 0.5))
grid.arrange(p1, p4,p2, p3, ncol = 4,
             top =textGrob("Distribution of important variables" ,
                          gp=gpar(fontsize=12,font=3))) # testosterone looks similar, except for scale.
```

### Variable selection

To choose what other variables to include in further analysis, we visually inspected pairwise correlations. We removed the following:

* time.of.day : didn't seem to have a straightforward relationship with other variables. 
* 'interim'variables: contain missing data for many teams.
* other 'final' variables besides final.performance: these variables are generally correlated, but not in a straightforward way. The original study states that these measures were standardized and then averaged to form the final.performance score. Since the judges who assigned the final.performance are also domain experts, we have decided to discard the other variables.

We then looked more closely at the variables 'females' and 'age' which may contribute to diversity outside of diversity score.

Gender is incorporated in the diversity score, but the variable females was measured separately. Since we are looking at average hormone scores which are known to differ between genders, we examined the proportion of females per group. As seen below, proportion of females shows little correlation with diversity score. However, there is  similar negative correlation with both testosterone and team.size. Our hypothesized causal graph suggested proportion of females should influence performance through both diversity and through affecting testosterone level. In contrast the similar pearson correlation coefficients suggest that proportion of females may not play a large role in determining performance outside of its effects on testosterone.

```{r females, fig.width=8, fig.height=6, echo=FALSE}
p1 <-ggplot(team_dat, aes(x = proportion.females)) +
  geom_histogram(bins =4)
p2 <-ggplot(team_dat, aes(x = proportion.females, y = score)) +
  geom_count()
# p3 <-ggplot(team_dat, aes(x = females/team.size, y = avg.log.testosterone)) +
#   geom_point()
p3 <-ggscatter(team_dat, x = "proportion.females", y = "avg.log.testosterone",
               cor.coef = TRUE,  add = "reg.line", cor.method = "pearson")
p4 <-ggscatter(team_dat, x = "proportion.females", y = "final.performance",
               cor.coef = TRUE,  add = "reg.line", cor.method = "pearson")
grid.arrange(p1, p2, p3, p4, ncol = 2,
             top =textGrob("Proportion of females is negatively correlated with both testosterone and performance",
                          gp=gpar(fontsize=12,font=3)))
#grid.arrange(p1, p2, p3, ncol = 3)
```

A similar situation occurs for age. When we plotted the spread of age across groups (below left), it seemed mean age is not too different between groups. Since according to our causal graph, the effects of age may influence performance through changing cooperation, we further plotted the variance of age since different ages in a group could also lead to conflict. It also doesn't correlate clearly with diversity score but does show a similar weak positive correlation with both testosterone and performance (see pearson correlation scores). This suggests that age may not play a large role in determining performance outside of its effects on testosterone.

```{r age, fig.width=8, fig.height=6, warning=FALSE, echo=FALSE}
p1 <-ggplot(ind_dat, aes(x= team.id, y=Age))+
  geom_boxplot()+theme_hw
p2 <-ggplot(team_dat, aes(x = age.variance, y = score)) +
  geom_count()
# p3 <-ggplot(team_dat, aes(x = avg.age, y = avg.log.testosterone)) +
#   geom_point()
p3 <-ggscatter(team_dat, x = "age.variance", y = "final.performance",
               cor.coef = TRUE,  add = "reg.line", cor.method = "pearson")
p4 <-ggscatter(team_dat, x = "age.variance", y = "avg.log.testosterone",
               cor.coef = TRUE,  add = "reg.line", cor.method = "pearson")
grid.arrange(p1, p2, p4, p3, ncol = 2,
             top =textGrob("Variance of age measures of diversity than diversity score",
                          gp=gpar(fontsize=12,font=3)))
#grid.arrange(p1, p2, p3, ncol = 3)
```

Based on the above, we decided to keep discard both age and proportion of females as variables since they may not necessarily tell us more about the relationship between testosterone and performance. Furthermore this simplifies the model.

## Results

Here we perform statistical analysis to verify our hypothesis. The results presented by the original study [@Akinola2018] include that:

* when group diversity was low, group testosterone significantly positively predicted performance at p < .01
* when group diversity was relatively high, group testosterone significantly negatively predicted performance p < .01

### Effect of diversity score and testosterone individually on performance
To start we want to check the simplest assumptions that 1) diversity score positively predicts performance and 2) testosterone negatively predicts performance. 

```{r model, fig.width=8, fig.height=4, warning=FALSE, echo=TRUE}
mod_t = lm( final.performance ~avg.log.testosterone, data = team_dat)
mod_d = lm( final.performance ~score, data = team_dat)
#summary(mod_d)
#summary(mod_t)
```
For these simple models, the coefficient of diversity score (0.065) and average log testosterone (0.7032) are not particularly large in magnitude or significant (p >0.05) indicating that neither of these two predictors alone shows a clear relationship with group performance. This agrees with what the authors found.

### Q4: Effect of group diversity on relationship between testosterone and performance
Next we fit a model to examine whether the interaction between them could predict performance. 

```{r test_model, fig.width=4, fig.height=4, warning=FALSE}
# model with normalized score
mod3=lm(final.performance ~ avg.log.testosterone + score + team.size + avg.log.testosterone:score, data = team_dat)
summary(mod3)
#plot(mod, which=1, col=c("blue"))
```
We found a significant positive effect of testosterone on performance when controlling for diversity score and team size (coefficient = 10.2743, p < 0.001). Further we find a significant positive effect of diversity score on performance when controlling for testosterone, team size and their interaction (coefficient = 10.1889, p < 0.001). The interaction term has a negative coefficient. This suggests that whereas each of testosterone and diversity aids performance, their interaction works against these effects. Our results are in line with those of the original study.

### Q5: Effect of cortisol on relationship between diversity and performance
We then examined whether stressed groups have better or worse performance by looking at pearson correlation. In the diagnostic plot, the correlation of cortisol with final.performance seems weaker than the correlation of testosterone with final.performance as shown below. However the r squared value shows that there is still some linearity in this relationship.

```{r horm, fig.width=8, fig.height=2.5, warning=FALSE}
t <- cor(team_dat$avg.log.testosterone,  team_dat$final.performance)   
c <- cor(team_dat$avg.log.cortisol,  team_dat$final.performance) 
p1 <-ggscatter(team_dat, x = "avg.log.testosterone", y = "final.performance",
               cor.coef = TRUE, add = "reg.line", cor.method = "pearson") 
p2 <-ggscatter(team_dat, x = "avg.log.cortisol", y = "final.performance",
               cor.coef = TRUE,  add = "reg.line", cor.method = "pearson")
grid.arrange(p1, p2, ncol = 2,
             top =textGrob("Relationship between hormone level and performance",
                          gp=gpar(fontsize=12,font=3)))
```

```{r}
mod_c <- lm(final.performance ~ avg.log.cortisol, data = team_dat)
# summary(mod_c)
```
When we fit the very simplest model of final.performance ~ avg.log.cortisol, we find a positive (0.1217) but not significant (p-value 0.56) coefficient as our scatterplots above may suggest.

### Model with interaction of cortisol and diversity score

Next, we tested whether cortisol levels could change the relationship between diversity score and performance with a model containing each of these variables and their three way interaction.

```{r stress, fig.width=8, fig.height=4, warning=FALSE}
#  will stress change the effect of diversity score on performance?
mod = lm(final.performance ~ avg.log.cortisol + score + team.size +avg.log.cortisol:score, data = team_dat)
summary(mod)
#plot(mod)
```
Here we found that stress seems to positively impact performance (coefficient of 3.348 units, p < 0.01) when controlling for diversity score, team size and the interaction between cortisol and diversity score. However, the diversity score is estimated here to have a negative effect on performance (coefficient of -6.8455 units, p < 0.05). Furthermore the interaction term also has a weak negative effect (coefficient of -0.7483 units, p < 0.05). 

This suggests stressed groups have better performance and stress changes the effect of diversity to negatively impact performance.

### Incorporating both hormone measurements

This suggests cortisol should also be in our model, so we lastly fit a full model with both hormone measurements and their three way interaction with diversity.

```{r complete, fig.width=8, fig.height=4, warning=FALSE}
#  will stress change the effect of diversity score on performance?
mod = lm(final.performance ~ avg.log.cortisol + avg.log.testosterone + score + team.size +avg.log.cortisol:score:avg.log.testosterone, data = team_dat)
summary(mod)
#plot(mod)
```
Here the coefficients on cortisol is positive (2.95) but the coefficient on diversity score is negative (-1.23) at p < 0.05, which is what we would expect. However the coefficient on testosterone is not significantly nonzero contrary to our hypothesis that group testosterone is important. Hoever the model could be a bit too complex since it has a lower R squared value than the model without cortisol.

### Conclusion

Here we have analyzed demographic data and hormone measurements from groups of MBA students performing a competetive project, previously published by [@Akinola2018]. We sought to investigate the authors' hypothesis that group diversity has a testosterone-dependent effect on group performance and also to check whether cortisol levels had an effect on this relationship.

By building linear models of performance, we have shown that although testosterone and diversity score alone do not predict performance, indeed when they are both included in the model interaction between diversity and testosterone has a significant negative effect on performance (p < 0.01) implying that high diversity and high testosterone are antagonizing factors. Although stressed groups did not have significantly different performance, we also found that when controlling for diversity cortisol has similar effects. The interaction between cortisol and diversity also has a significant negative effect on performance (p < 0.05) implying that higher diversity and higher cortisol counteract each other. When looking at both hormone measurements simultaneously with diversity score, surprisingly we found that when accounting for cortisol, testosterone levels do not seem to have a significant effect on performance. Rather only the interaction of cortisol and testosterone together has a slight negative effect on performance (p < 0.01). Overall, this suggests that perhaps, stress has a more important underlying role in group performance than previously suggested and may mediate effects seen in the relationship between testosterone and performance.

Although we had some similar findings to the original study when examining diversity and testosterone, our results may not be directly comparable because of some differences in our methodology. Most prominently, they have used a faultline analysis to evaluate diversity whereas we have constructed a diversity score. As well, we have not included some of the variables that are present in the models which they tested e.g. proportion of females. We chose to discard these variables based upon our EDA and our reasoning about the relationship between variables collected in the study. Lastly we cannot compare our findings about cortisol because this was not discussed in depth in their original analysis.

# Bibliography
