---
title: "Hw6_q3"
date: "10/7/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, width.cutoff=60)

# Weisberg, Applied Linear Regression, 3rd ed., question 12.1 (pp. 266–267). You can obtain the
# data as the downer variable in the package alr3.
# You will have to remove cases for which observations are missing whenever you need those
# observations to fit a model.
# In 12.1.3, your comments should address what the two density plots (you can make those using
# ordinary kernel density smoothing) imply about how easy it would be to predict survival using
# this variable.
library(alr3)
# libraries
library(tidyverse)
library(lattice)
library(dagitty)
library(reshape2)
library(grid)
library(gridExtra)
library(np)
library(stargazer)
library(knitr)

suppressWarnings(suppressMessages(library(stargazer))) 
knitr::opts_chunk$set(warning=FALSE,width.cutoff=60)
head(downer)
```

## 1. 

Myopathy is columns, and cows outcome is on teh y. We see that the percent of surviving cow (outcome =1) is 39% for those without myopathy and 6% for those with myopathy.

```{r cars}
prop.table(table(downer$Outcome, downer$Myopathy), 2)
```

## 2. Fit logistic regression.

Based on the coefficient table the estimated log probability when myopathy = 0 would be -0.4649 ± 0.1823, and when myopathy = 1 would be -2.2320 ± 0.4595. The decrease in odds of survival from mypathy is then exp(-2.2320)=0.11. Then the probability of survival when myopathy=1 is 0.11*exp(-0.4649) = 0.07, and when myopathy=0 is 1-exp(-0.4649) = 0.37. These numbers are about where the last row from Q1 are.

```{r pressure, echo=FALSE}
library(MASS)
fit <- glm(Outcome ~ Myopathy, 
           data=downer, family=binomial(link="logit"))
summary(fit)
```
## 3.

We plotted the density of the CK and log(CK) below. It seems like the dead cows had higher levels of phosphokinase, and this is a bit clearer to see when we have log scale x-axis. Also the values look more normally distributed in the second graph.

```{r }
library(sm)
temp <- downer[, c("Outcome", "CK")] %>% na.omit()
sm.density.compare( temp$CK, as.factor(temp$Outcome))
title("Serum creatine phosphokinase")
colfill<-c(2:3)
legend("right", fill=colfill, 
       legend=c("Died", "Survived"))

graphics.off()

sm.density.compare(log(temp$CK), as.factor(temp$Outcome))
title("Serum creatine phosphokinase")
colfill<-c(2:3)
legend("right", fill=colfill, 
       legend=c("Died", "Survived"))
```


## 4

Here are the results, showign that the fold increase in survival for each unit increase in log(CK) is exp(-0.6117)=0.542428. This suggests perhaps myopathy is more predictive.

```{r, echo=FALSE}
fit <- glm(Outcome ~ log(CK), 
           data=downer, family=binomial(link="logit"))
summary(fit)
```


## 5

Here are the results of the model:
```{r, echo=FALSE}
temp <- downer[, c("Outcome", "CK", "Myopathy")] %>% na.omit()
fit <- glm(Outcome ~ log(CK) + Myopathy + log(CK):Myopathy, 
           data=downer, family=binomial(link="logit"))
summary(fit)
```

Additionally we can show the deviance per term:
```{r}
anova(fit, test = "Chisq")
```

It seems like the effect of each of log(CK) and Myopathy is significant (p < 0.001) but their interaction is not.