---
title: "Kernel Smoothing and Densities"
author: "Alex Reinhart"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kernel Smoothing

For the ordinary Nadaraya--Watson kernel smoother, R's built in `ksmooth` function works fine:

```{r}
plot(cars$speed, cars$dist)
lines(ksmooth(cars$speed, cars$dist, kernel="normal", bandwidth=1))
lines(ksmooth(cars$speed, cars$dist, kernel="normal", bandwidth=2), col=2)
lines(ksmooth(cars$speed, cars$dist, kernel="normal", bandwidth=3), col=3)
```

`ksmooth` is very simple. You can pick a couple different kernels and can specify the bandwidth, but it does not do local linear or local polynomial kernel regression, as ESL describes in 6.1.1.

The `np` package from CRAN is a very flexible package for nonparametric regression. It supports many kernels, has methods for selecting bandwidths, and can do regressions with many columns in X, including discrete variables. We run `npregbw` to let it select bandwidths for each X variable automatically (using cross-validation methods), then provide these bandwidths to `npreg`:

```{r}
suppressMessages(library(np))

## "ll" for "local-linear"; "lc" does Nadaraya--Watson
bw <- npregbw(dist ~ speed, data=cars, regtype="ll")
bw

loclinfit <- npreg(bw)
loclinfit

plot(loclinfit)
points(cars$speed, cars$dist)
```

You can see the best cross-validated bandwidth is somewhat wider than the ones we were trying for `ksmooth`.

`npreg` supports continuous data, categorical data, and even ordered variables.

## Kernel Density Estimation

The built-in function `density` does univariate kernel density estimation. It takes a vector of points and returns a list with `x` points and the densities (`y`) corresponding to each point. It lets you select a bandwidth or uses several simple rules to pick a bandwidth: it can use cross-validation, Silverman's rule (setting the bandwidth proportional to the standard deviation in a complicated way), or a few other heuristic methods.

```{r}
plot(density(cars$speed, bw="ucv"))
rug(cars$speed)
```

Want a fancier kernel density estimator for multiple variables? Consider the `npudens` function from the `np` package. You first use `npudensbw` to select a bandwidth (using cross-validation or other rules), then use `npudens` to get a density estimate. For example, we can get the joint density of `speed` and `dist`:

```{r}
bw <- npudensbw(~ speed + dist, data=cars)

dens <- npudens(bw)
plot(dens) ## fancy animation -- doesn't show up in Rmd well
```