---
title: "Regression diagnostic plots"
author: "Prof. Chouldechova"
date: ""
output:
  html_document:
    toc: true
    toc_depth: 5
---

```{r}
library(ggplot2)
```

### Introduction

This set of supplementary notes provides further discussion of the diagnostic plots that are output in R when you run th `plot()` function on a linear model (`lm`) object.  

### 1. Residual vs. Fitted plot


#### The ideal case

Let's begin by looking at the Residual-Fitted plot coming from a linear model that is fit to data that perfectly satisfies all the of the standard assumptions of linear regression.  That is, we'll generate data according to 

$$y_i = 3 + 0.1 x + \epsilon_i,$$ 

for   $i = 1, 2, \ldots, 1000$, where the $\epsilon_i$ are independent Normal$(0,sd = 3)$ variables (with standard deviation 3).  

Here's code to generate this data and then regress y on x.

```{r}
n <- 1000      # sample size
x <- runif(n, min = 0, max = 100)
y.good <- 3 + 0.1 * x + rnorm(n, sd = 3)

# Scatterplot of the data with regression line overlaid
qplot(x, y.good, ylab = "y", main = "Ideal regression setup") + stat_smooth(method = "lm")

# Run regression and display residual-fitted plot
lm.good <- lm(y.good ~ x)
plot(lm.good, which = 1)  
```

The scatterplot shows the perfect setup for a linear regression:  The data appear to be well modeled by a linear relationship between $y$ and $x$, and the points appear to be randomly spread out about the line, with no discerninle non-linear trends or changes in variability.  

Looking at the Residuals vs Fitted plot, we see that the red line (which is just a scatterplot smoother, showing the average value of the residuals at each value of fitted value) is perfectly flat.  This tells us that there is no discernible non-linear trend to the residuals.  Furthermore, the residuals appear to be equally variable across the entire range of fitted values.  There is no indication of non-constant variance.

#### Curvature or non-linear trends

Here's an example where we have non-linear trends in the data.  This example is constructed to mimic seasonal data.  

```{r, echo=FALSE}
y.curved <- 5 * sin(0.6 * x) + 0.1 * x + rnorm(n, sd = 2)

# Scatterplot of the data
qplot(x, y.curved, ylab = "y", main = "Data with a seasonal trend") + 
  stat_smooth(method = "lm") + stat_smooth(method = "loess", span = 0.1, colour = I("red"))
```

The blue line shows the model fit.  The red curve is a non-linear fit that does a better job of modelling the average value of $y$ at each value of $x$.   Note that the linear model fails to capture the clear non-linear trend that's present in the data.  This causes tremendous problems for our inference.  Look at the grey confidence band that surrounds the regression line.  If the standard linear regression assumptions are satisfied, this band with high likelihood would contain the average value of $y$ at each value of $x$.  i.e., the grey bands around the blue curve should mostly contain the red curve.  This obviously does not happen.  The red curve is almost always far outside the grey bands around the blue regression line.  

**Take-away**:  When one or more of the model assumptions underlying the linear model is violated, we can no longer believe our inferential procedures.  E.g., our confidence intervals and p-values may no longer be reliable.

Here's what the Residual - Fitted plot looks like for this model.

```{r}
lm.curved <- lm(y.curved ~ x)

plot(lm.curved, which = 1)
```

Visually, we see a clear trend in the residuals.  They have a periodic trend.  Unfortunately, the scatterplot smoother that's used to construct the red line isn't doing a good job here.  This is a case where the choice of neighbourhood size (how many points go into calculating the local average) is taken to be too large to capture the the trend that we visually observe.  Don't always trust that red curve.  

##### Constructing your own Residual vs Fitted plot

Here's a better version of the default plot.  

```{r}
# Plot model residuals on y axis, fitted values on x axis
# Add red trend curve with better choice of smoothing bandwidth
qplot(y = lm.curved$residuals, x = lm.curved$fitted.values,
      ylab = "Residuals", xlab = "Fitted values", 
      main = "The Do-it-yourself Residuals vs. Fitted plot") +
  stat_smooth(method = "loess", span = 0.1, colour = I("red"), se = FALSE)
```

#### Non-constant variance

```{r}
y.increasing <- 3 + 0.2 * x + (1 + x / 25) * rnorm(n, sd = 3)

# Produce scatterplot of y vs x
qplot(x, y.increasing, ylab = "y")
```

Here's what the Residual vs. Fitted plot looks like in this case.

```{r}
lm.increasing <- lm(y.increasing ~ x)
plot(lm.increasing, which = 1)
```

If you look at this plot, you'll see that there's a clear "funneling" phenomenon.  The distribution of the residuals is quite well concentrated around 0 for small fitted values, but they get more and more spread out as the fitted values increase. This is an instance of "increasing variance".  The standard linear regression assumption is that the variance is constant across the entire range.  When this assumption isn't valid, such as in this example, we shouldn't believe our confidence intervals, prediction bands, or the p-values in our regression.  


### Normal QQ plot

The Normal QQ plot helps us to assess whether the residuals are roughly normally distributed.  If the residuals look far from normal we may be in trouble.  In particular, if the residual tend to be larger in magnitude than what we would expect from the normal distribution, then our p-values and confidence intervals may be too optimisitic.  i.e., we may fail to adequately account for the full variability of the data.


#### The ideal case

First, here's an example of a Normal QQ plot that's as perfect as it gets.  This comes from the ideal simulation setting in the previous section.  The residuals here are a perfect match to the diagonal line.  These residuals look to be normally distributed.

```{r}
plot(lm.good, which = 2)
```

#### Lighter tails

In the next example, we see a QQ plot where the residuals deviate from the diagonal line in both the upper and lower tail.  This plot indicated that the tails are 'lighter' (have smaller values) than what we would expect under the standard modeling assumptions.  This is indicated by the points forming a "flatter" line than than the diagonal.

```{r}
plot(lm.curved, which = 2)
```

#### Heavier tails

In this final example, we see a QQ plot where the residuals deviate from the diagonal line in both the upper and lower tail.  Unlike the previous plot, in this case we see that the tails are observed to be 'heavier' (have larger values) than what we would expect under the standard modeling assumptions.  This is indicated by the points forming a "steeper" line than the diagonal.

```{r}
plot(lm.increasing, which = 2)
```

### Outliers and the Residuals vs Leverage plot

There's no single accepted definition for what consitutes an outlier.  One possible definition is that an outlier is any point that isn't approximated  well by the model (has a large residual) and which significantly influences model fit (has large leverage).  This is where the Residuals vs Leverage plot comes in.  

#### The ideal case

Let's look at our ideal setting once again.  The plot below is a great example of a Residuals vs Leverage plot in which we see no evidence of outliers. Those "Cook's distance" dashed curves don't even appear on the plot.  None of the points come close to having both high residual and leverage.

```{r}
plot(lm.good, which = 5)
```

#### An example with possible outliers

```{r}
set.seed(12345)
y.corrupted <- y.good[1:100]
x.corrupted <- x[1:100]

# Randomly select 10 points to corrupt
to.corrupt <- sample(1:length(x.corrupted), 10)
y.corrupted[to.corrupt] <- - 1.5 * y.corrupted[to.corrupt] + 3 * rt(10, df = 3)
x.corrupted[to.corrupt] <- x.corrupted[to.corrupt] * 2.5
# Fit regression and display diagnostic plot
lm.corrupted <- lm(y.corrupted ~ x.corrupted)

plot(lm.corrupted, which = 5)
```

In this plot we see that there are several points that have high residual and high leverage.  The points that lie close to or outside of the dashed red curves are worth investigating further.  

### Can't we just use scatterplots?

All of the examples above were generated by considering the regression of a single outcome variable on a single covariate.  In this case, we could've diagnosed most of the violations of model assumptions just by looking at the x-y scatterplot.  The reason for using diagnostic plots is that most regressions we run aren't so simple.  Most regressions use many variables (tens or hundreds of variables), and in those cases there isn't a good way of visualizing all of the data.  Residuals, fitted values and leverage are all quantities that can be computed and plotted regardless of how many variables are included in the model.  Thus diagnostics such as the Residual vs. Fitted plot, Normal QQ plot and Residual vs. Leverage plot can help us even when we have complicated models.  